---
title: "homework"
author: Isabella Muratore
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

# homework 02

## What is the probability that she will hear more than 8 calls during any given session?
### using a poisson probability distribution to get the upper tail using 9 position along the distribution because that's how I interpretted "more than 8"

```{r}

ppois(9, lambda=15, lower=FALSE) 

```

#### returns 0.93 etc, so the answer is a ~93% probability of more than 8 assuming the mean is accurate

## What is the probability that she will hear no calls in a session?
### using a poisson probability distribution to get the lower tail using 0 as the position along the distribution

```{r}

ppois(0, lambda=15)

```

#### returns 0.0000003059023, so a 0.00003059023% probability, quite unlikely

## What is the probability that she will hear exactly 3 calls in a session?
### using a poisson density distribution (since this is just for one specific value) to get the probability of hearing exactly 3 calls

```{r}

dpois(3, lambda=15)

```

#### returns 0.0001720701, so 0.01720701% probability, also unlikely

## Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30
### Using dpois from 0 to 30 and plotting as points over x = a vector from 0 to 30

```{r}

z <- dpois(0:30, lambda=15)

x <- c(0:30)

barplot(z, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = "Poisson Mass Function")

```

####

## Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
### 

```{r}

x <- rpois(104, lambda=15)

#ignore
#set <- NULL  # sets up a dummy variable to hold our 10000 simulations
#n <- 104
#for (i in 1:104) {
#    set[i] <- mean(sample(x, n, replace = TRUE))
#}

#set

x

```

## Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?

### applying hist() with an x range of 0-30 to "set", which we filled using the bootstrapping loop above

```{r}

hist(x, xlim=range(c(0:30)))



#### This distribution is much narrower than the one made using the probability mass function. This happens because increasing the sample size or number of simulated values tightens the distribution and decreases variance/standard deviation because n is in the denominator of these equations.

```

#### This distribution is much less smooth and symmetrical than the one made using the probability mass function. The probability mass function histogram is more "perfect" because it has an infinite sample size, as opposed to 104, and this tightens the distribution and decreases variance/standard deviation because n is in the denominator of these equations.

# second homework problem
## Load in the dataset “zombies.csv” from my GitHub repo
### downloading zombies.csv from GitHub using curl, formatting using read.csv with commas as separation

```{r}
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/zombies.csv")

d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)

```
## Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and  sd() commands as these are for samples.
### using mean function to average the height variable

```{r}

h <- d$height

mean(h)

w <- d$weight

mean(w)

a <- d$age

mean(a)

z <- d$zombies_killed

mean(z)

y <- d$years_of_education

mean(y)

```
###

```{r}
pop_v <- function(x) {
    sum((x - mean(x))^2)/(length(x))
}
pop_v(h)

pop_v(w)

pop_v(a)

pop_v(z)

pop_v(y)

```
###

```{r}

pop_sd <- function(x) {
    sqrt(pop_v(x))
}
pop_sd(h)

pop_sd(w)

pop_sd(a)

pop_sd(z)

pop_sd(y)

```

