---
title: "homework"
author: Isabella Muratore
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

# homework 02

## What is the probability that she will hear more than 8 calls during any given session?
### using a poisson probability distribution to get the upper tail using 8 position along the distribution.

```{r}

ppois(8, lambda=15, lower=FALSE) 

```

#### returns ~0.96, so the answer is a ~96% probability of more than 8 assuming the mean is accurate

## What is the probability that she will hear no calls in a session?
### using a poisson density distribution (since this is just for one specific value) to get the probability of hearing exactly 0 calls (you could also do this with ppois at position 0 capturing the lower tail)

```{r}

dpois(0, lambda=15)

```

#### returns 0.0000003059023, so a 0.00003059023% probability, quite unlikely

## What is the probability that she will hear exactly 3 calls in a session?
### using a poisson density distribution (since this is just for one specific value) to get the probability of hearing exactly 3 calls

```{r}

dpois(3, lambda=15)

```

#### returns 0.0001720701, so 0.01720701% probability, also unlikely

## Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30
### Using dpois from 0 to 30 and plotting as points over x = a vector from 0 to 30

```{r}

z <- dpois(0:30, lambda=15)

x <- c(0:30)

barplot(z, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = "Poisson Mass Function Modeling Monkey Calls")

```

#### this creates a smooth looking distribution

## Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
### Using rpois to generate 104 random numbers from a poisson distribution with lambda 15

```{r}

x <- rpois(104, lambda=15)

x

```
#### returns a vector of numbers

## Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?
### applying hist() with an x range of 0-30 to "x", which we filled with random poisson numbers above

```{r}

hist(x, xlim=range(c(0:30)), xlab=c(0:30), main="Poisson Mass Function Modeling Monkey Calls")

```

#### This distribution is much less smooth and symmetrical than the one made using the probability mass function. The probability mass function histogram is more "perfect" because it represents the entire infinitely divisible distribution, as opposed to showing 104 discrete elements from the distribution, and larger sample size tightens the distribution and decreases variance/standard deviation because n is in the denominator of these equations.
