---
title: "homework3"
author: "Isabella Muratore"
date: "September 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# second homework problem

## Load in the dataset “zombies.csv” from my GitHub repo
### downloading zombies.csv from GitHub using curl, formatting using read.csv with commas as separation

```{r}
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/zombies.csv")

d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)

```
## Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and  sd() commands as these are for samples.
### using mean function to average each of the variables from the df

```{r}

h <- d$height

mean(h)

w <- d$weight

mean(w)

a <- d$age

mean(a)

z <- d$zombies_killed

mean(z)

y <- d$years_of_education

mean(y)

```
#### means for height, weight, age, zombies killed, years of education

### a function for population variance summing the average deviation from the mean

```{r}
pop_v <- function(x) {
    sum((x - mean(x))^2)/(length(x))
}


```

### a function for taking the square root of population variance to generate population standard deviation, calculating standard deviation for each of the variables from the df

```{r}

pop_sd <- function(x) {
    sqrt(pop_v(x))
}
pop_sd(h)

pop_sd(w)

pop_sd(a)

pop_sd(z)

pop_sd(y)

```
## Use {ggplot} and make boxplots of each of these variables by gender.
###
```{r}
library(ggplot2)

p <- ggplot(data = d, aes(x = gender, y = h))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("height")
p

p <- ggplot(data = d, aes(x = gender, y = w))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("weight")
p

p <- ggplot(data = d, aes(x = gender, y = a))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("age")
p

p <- ggplot(data = d, aes(x = gender, y = z))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("zombies killed")
p

p <- ggplot(data = d, aes(x = gender, y = y))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("years of education")
p



```

##Use {ggplot} and make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?
###

```{r}

p <- ggplot(data = d, aes(x = age, y = height))  # first, we build a plot object
p <- p + xlab("age") + ylab("height")  # then we modify the axis labels
p <- p + geom_point()  # then we make a scatterplot
p <- p + theme(legend.position = "bottom", legend.title = element_blank())  # then we modify the legend
p  # and, finally, we plot the object

p <- ggplot(data = d, aes(x = age, y = weight))  # first, we build a plot object
p <- p + xlab("age") + ylab("weight")  # then we modify the axis labels
p <- p + geom_point()  # then we make a scatterplot
p <- p + theme(legend.position = "bottom", legend.title = element_blank())  # then we modify the legend
p  # and, finally, we plot the object

```

#### Without quantifying correlation, I would say that both height and weight are positively correlated to age, with the correlation between weight and age appearing slightly less robust than that between height and age.

## Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not (hint: not all are drawn from the normal distribution)? For those that are not, can you determine what common distribution they are drawn from?
```{r}
hist(h)

qqnorm(h, main = "Height Normal Q-Q Plot")
qqline(h, col = "gray")
#normal

hist(w)

qqnorm(w, main = "Weight Normal Q-Q Plot")
qqline(w, col = "gray")
#normal

hist(a)

qqnorm(a, main = "Age Normal Q-Q Plot")
qqline(a, col = "gray")
#normal

hist(z)

qqnorm(z, main = "Zombies Killed Normal Q-Q Plot")
qqline(z, col = "gray")
#poisson 

hist(y)

qqnorm(y, main = "Education Normal Q-Q Plot")
qqline(y, col = "gray")
#poisson

```
####

## Now use the sample() function to sample ONE subset of 30 zombies (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you will need to base your estimate of the CIs on some different distribution.
### For each variable I am creating a sample "s" that is a vector of one sample of thirty values for height, or weight, or age, etc. drawn without replacement from the original population.  

```{r}
k <- 1  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
    s[[i]] <- sample(h, size = n, replace = FALSE)
}
```
```{r}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  
```
```{r}
stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
sem <- stdev/sqrt(n)  # a vector of SEs estimated from each sample 
```
```{r}
upper <- mean(m) + qnorm(0.95, mean = 0, sd = 1) * mean(sem)
lower <- mean(m) + qnorm(0.05, mean = 0, sd = 1) * mean(sem)
ci <- c(lower, upper)
ci
```
#### Now repeating this process for the other variables that have normal distributions
```{r}
k <- 1  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
  #for weight
    s[[i]] <- sample(w, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
sem <- stdev/sqrt(n)  # a vector of SEs estimated from each sample 

upper <- mean(m) + qnorm(0.95, mean = 0, sd = 1) * mean(sem)
lower <- mean(m) + qnorm(0.05, mean = 0, sd = 1) * mean(sem)  
ci <- c(lower, upper)
ci

####

k <- 1  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
  #for age
    s[[i]] <- sample(a, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
sem <- stdev/sqrt(n)  # a vector of SEs estimated from each sample 

upper <- mean(m) + qnorm(0.95, mean = 0, sd = 1) * mean(sem)
lower <- mean(m) + qnorm(0.05, mean = 0, sd = 1) * mean(sem)  
ci <- c(lower, upper)
ci

####

k <- 1  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
  # for zombies killed
    s[[i]] <- sample(z, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
sem <- stdev/sqrt(n)  # a vector of SEs estimated from each sample 

#upper <- mean(m) + qpois(0.95, lambda=mean(m), lower.tail = #TRUE, log.p = FALSE) * mean(sem)
#lower <- mean(m) + qpois(0.05, lambda=mean(m), lower.tail = #TRUE, log.p = FALSE) * mean(sem) 

upper <- mean(m)+1.96*sqrt(mean(m)/n)
lower <- mean(m)-1.96*sqrt(mean(m)/n)
ci <- c(lower, upper)
ci

```
#### Now using the same tools to sample from the remaining two variables and obtain means and standard deviations, but using a different formula to get poisson-based confidence intervals. The upper bound is obtained by adding the product of 1.96 and the square root of lambda over n to lambda; the lower bound is obtained by subtracting this value from lambda. 

```{r}

k <- 1  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
  # for years of education
    s[[i]] <- sample(y, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
sem <- stdev/sqrt(n)  # a vector of SEs estimated from each sample 

#upper <- mean(m) + qpois(0.95, lambda=mean(m), lower.tail = #TRUE, log.p = FALSE) * mean(sem)
#lower <- mean(m) + qpois(0.05, lambda=mean(m), lower.tail = #TRUE, log.p = FALSE) * mean(sem)  

upper <- mean(m)+1.96*sqrt(mean(m)/n)
lower <- mean(m)-1.96*sqrt(mean(m)/n)

ci <- c(lower, upper)
ci



```
####

## Now draw 99 more random samples of 30 zombies out and calculate the mean for each of the these samples. Together with the first sample you drew out, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of this distribution for each variable? How do the standard deviations compare to the standard errors estimated in [5]? What do these sampling distributions look like? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?
###

```{r}
k <- 99  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
    s[[i]] <- sample(h, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  
hist(m, main = "Height Sample Histogram")
qqnorm(m, main = "Height Sample Normal Q-Q Plot")
qqline(m, col = "gray")

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
mean(stdev)

k <- 99  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
    s[[i]] <- sample(w, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  
hist(m, main = "Weight Sample Histogram")
qqnorm(m, main = "Weight Sample Normal Q-Q Plot")
qqline(m, col = "gray")

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
mean(stdev)

k <- 99  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
    s[[i]] <- sample(a, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  
hist(m, main = "Age Sample Histogram")
qqnorm(m, main = "Age Sample Normal Q-Q Plot")
qqline(m, col = "gray")

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
mean(stdev)

k <- 99  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
    s[[i]] <- sample(z, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  
hist(m, main = "Zombies Killed Sample Histogram")
qqnorm(m, main = "Zombies Killed Sample Normal Q-Q Plot")
qqline(m, col = "gray")

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
} 
mean(stdev)

k <- 99  # number of samples
n <- 30  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
    s[[i]] <- sample(y, size = n, replace = FALSE)
}

m <- NULL
for (i in 1:k) {
    m[i] <- mean(s[[i]])
}
mean(m)  
hist(m, main = "Education Sample Histogram")
qqnorm(m, main = "Education Sample Normal Q-Q Plot")
qqline(m, col = "gray")

stdev <- NULL
for (i in 1:k) {
    stdev[i] <- sd(s[[i]])
}
mean(stdev)

```
#### Interestingly, both of the "poisson distributed" variables (zombies killed and years of education) appear fairly normally distributed based on the histograms and qqplots generated from random samples of their values. This is likely due to the central limit theorem which states, according to Module 9, "the distribution of averages (or sums or other summary statistics…) of iid (independent and identically distributed) random variables becomes normal as the sample size increases... no matter what kind of probability distribution characterizes the initial population"





